---
title: "Homework4"
author: "Tyler Poelking"
date: "2/13/2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

Question 1)

a.	A non-informative prior would be if θ ~ Uniform(-1, 1). This gives θ  an equal chance to be anywhere between (-1, 1). I chose these parameter values because the difference in shots made frequency cannot be less than -1 nor greater than 1.
b.	A subjective prior based on my knowledge of free throw shooting would be θ ~ Normal(mean = 0.2, sd = 0.3). I chose 0.2 for the mean because it is likely that daily 50 practice shots over a month increases their shooting percentage but not by a drastic amount. I chose 0.2 because the percent improvement will vary a decent amount given student's prior skillset and their natural athletic abilities. (0.2-2(0.3), 0.2+2(0.3)) = (-0.4, 0.5).


Question 2)

a) Plot the prior density function for θ in R.
```{r Question 2 Part a}
betaplot <- function(a,b){
    theta = seq(0,1,0.005)
    p_theta = dbeta(theta, a, b)
    p <- qplot(theta, p_theta, geom='line')
    p <- p + theme_bw()
    p <- p + ylab(expression(paste('p(',theta,')', sep = '')))
    p <- p + xlab(expression(theta))
    return(p)}

sprintf("Beta Distribution with a = %s and b = %s", 4, 4)
betaplot(4,4)
```


b) Flip the coin one time and find the posterior distribution for θθ conditional on the outcome of the flip. Plot the posterior density function in R. 

In this case a head was received.
```{r Question 2 Part b}

a = 4
b = 4
# calculate posterior a
a2 = a + 1    
      				
# calculate posterior b
b2 = b + 1 -1

betaplot(a2,b2)


```


C) Now assume that the prior distribution for θ is the posterior distribution from part b. Flip the coin again. Find the the new posterior of θ and plot the posterior density function.

In this case a head was received again.
```{r Question 2 Part c}

# calculate posterior a
a3 = a2 + 1    
      				
# calculate posterior b
b3 = b2 + 1 -1

betaplot(a3,b3)
```


D) Assuming the original Beta(4,4) prior distribution for θθ, find the posterior distribution of θ conditional on the outcome of both coin flips. Compare this posterior distribution to the one you obtained in part c.
```{r Question 2 Part d}

# calculate posterior a
a4 = a + 2  
      				
# calculate posterior b
b4 = b + 2 -2

betaplot(a4,b4)
```

The two distributions are exactly the same, with a = 6 and b = 4. In part b we added 1 to 'a' and in part c we added 1 again. In part d we added 1 twice. In part b and c we added 1 and subtracted 1 from 'b'. In part d we added 2 and subtracted 2. 
