---
title: "Homework6_TylerPoelking"
author: "Tyler Poelking"
date: "3/6/2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rjags)
library(ggplot2)
library(grid)
library(gridExtra)
library(reshape2)
library(scales)
```

```{r Model Creation}
salaries <- read.table("salaries.txt",header=T)
#get salaries column
sal <- salaries$Salary
#n is total number entries
n <- length(sal)

#get unique number of departments. should be 8
dept <- salaries$Dept
deptNames <- levels( dept )
deptNumeric <- as.numeric( dept )

depts <- as.factor(salaries$Dept)
deptsUniqCount <- length(unique(depts))

#convert to numeric and extract counts of each dept
deptsNum <- as.numeric(depts)
deptsCounts <- as.vector(table(deptsNum))

#create empty matrix to hold our data. this is in a format such that each col corresponds to a dept. Note: due to varying
#salary amounts for each dept, not all data will be filled in. Regardless, this makes it easier for jags to process properly
sal.Matrix <- matrix(c(NA), ncol=deptsUniqCount,nrow=max(deptsCounts))
for (i in 1:deptsUniqCount){
  temp <- sal[which(deptsNum == i)]
  col <- c(temp, rep(NA, nrow(sal.Matrix)-length(temp)))
  sal.Matrix[,i] <- col
}

# create a data list
dataList = list('sal.Matrix' = sal.Matrix,  'deptsCounts' = deptsCounts, 'deptsUniqCount' =deptsUniqCount)


# list of parameters to be monitored  
parameters <- c("theta", "mu", "prec.within", "prec.btw")



# set initial values
initsValues <- list("theta" = rep(80,deptsUniqCount), "mu"=80, "prec.within" = 1/100, "prec.btw"=1/100)

# number of iteration for "tuning" 
adaptSteps <- 5000 

# number of iterations for "burn-in" 
burnInSteps <- 5000   

# number of chains to run
nChains <- 2          

# total number of iterations to save
numSavedSteps <- 5000           

# "thinning" (1 = keep every interation)
thinSteps <- 1                  

# iterations per chain
ITER <- ceiling((numSavedSteps * thinSteps )/ nChains) 

# -------------
# Run JAGS
# -------------

# create, initialize, and adapt the model
jagsModel <- jags.model("salaries_model.txt", data = dataList, inits = initsValues, n.chains = nChains, n.adapt = adaptSteps)

# burn-in the algorithm
update(jagsModel, n.iter = burnInSteps)

# run algorithm to get interations for inference
codaSamples <- coda.samples(jagsModel, variable.names = parameters, n.iter = ITER, thin = thinSteps)


```

Model Initial Values 
```{r}
initsValues
```

Model Initial Values and the  number of steps for tuning, burn-in, and for each chain were:
```{r}
initsValues

print(c(adaptSteps, burnInSteps, ITER))
```


Checking for model convergence and trace plot creation
```{r convergence + trace plots}

# -------------
# Look at posterior samples
# -------------

# make a dataframe with the posterior samples
mcmcChainDF <- data.frame(as.matrix(codaSamples, iters = T, chains = T))

# create a vector with the variable names
varNames <- names(mcmcChainDF)[3:(dim(mcmcChainDF)[2])]

# number of variables
nVars <- length(varNames)

mcmcChainDF$CHAIN <- as.factor(mcmcChainDF$CHAIN)
# construct trace plots
par(ask = T)
  for( i in 1:nVars ){
    print(ggplot(mcmcChainDF, aes( x = ITER, y = mcmcChainDF[ ,varNames[i]])) +
            geom_line(aes(color = CHAIN)) + 
            labs(y = varNames[i]))
    flush.console()
  }

```

The trace plots above shows no evidence against converge because the trace plots are consitent around one value and no pattern seems to be present that has not repeated. 




```{r Boxplots to summarize the posterior distribution of the components of θ and μ. }



postDFreshape <- melt( mcmcChainDF, id.vars = "ITER", measure.vars = c("theta.1.", "theta.2.", "theta.3.", "theta.4.", "theta.5.", "theta.6.", "theta.7.", "theta.8.", "mu"))

cols <- hue_pal()(4)

plot = ggplot(postDFreshape, 
       aes(x = variable, y = value )) +
     geom_boxplot( fill = c(rep(cols[3], deptsUniqCount), 
                          cols[1]),
                color = "darkgrey") +
    ggtitle("Posterior Distributions of Theta and Mu Parameters") + 
    ylab( "posterior" ) +
    xlab( "theta" ) +
    scale_x_discrete( labels = c("BM","CH", "CS", "CV", "EE", "IN", "MC", "MS", "mu" )) 

plot


```
At around 136k, the Biomedical department has the largest average salary of the 8 departments. Chemical comes in a fairly close second at just above 130k. Mechanical is the lowest at around 110k. The Average is about 120k dollars per year. 


Numerically summarize how the variation in salaries within discipline compares to the total variation in salaries. Include a numerical summary of the uncertainty associated with your inference about the sources of variation in salaries

```{r}
#use prec.within and prec.btw to calculate:
#sigma^2 - variance within samples 
#tau^2 - variance between a sample:
mcmcChainDF$sigma2 <- 1/mcmcChainDF$prec.btw
mcmcChainDF$tau2 <- 1/mcmcChainDF$prec.within

mcmcChainDF$propVar <- mcmcChainDF$sigma2 /
                        (mcmcChainDF$sigma2 +
                        mcmcChainDF$tau2)
```


We can find distributions of sigma^2 and tau^2, as well as the proportion of variance in the posterior distributions that is due to between-batch variance:

```{r}
par(ask = F)
  ggplot( mcmcChainDF, aes( x = sigma2 )) +
    geom_density( color = "red") +
    geom_histogram( aes(y = ..density.. ),
                    alpha = .3) 
  
mean( mcmcChainDF$sigma2 )
sd( mcmcChainDF$sigma2 )

par(ask = F)
  ggplot( mcmcChainDF, aes( x = tau2 )) +
    geom_density( color = "red") +
    geom_histogram( aes(y = ..density.. ),
                    alpha = .3) 
  
mean( mcmcChainDF$tau2 )
sd( mcmcChainDF$tau2 )

par(ask = F)
  ggplot( mcmcChainDF, aes( x = propVar )) +
    geom_density( color = "red") +
    geom_histogram( aes(y = ..density.. ),
                    alpha = .3) 
  
mean( mcmcChainDF$propVar)
sd( mcmcChainDF$propVar)
```
The expected proportion of variation explained by salaries within discipline is .177. This means that on average 17.7% of the variance is explained by salaries within discipline while the rest of the variance is explained by the salaries between disciplines. The standard deviation of the proportions is 0.072, which provides an idea on the variablility we expect to see in the proportion of variability within discipline salary is covers (lower sd implying higher confidence). 


What is the posterior probability that a new biomedical engineering faculty member makes more than $140K?
```{r}
bm.post <- mcmcChainDF$theta.1.

num_bm = length(bm.post)
probability <- length(which(bm.post > 140))/num_bm
print(probability)
```
