---
title: "hw 2 3303"
author: "Tyler Poelking"
date: "1/22/2018"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```
Problem 1: See Handwritten Page

Problem 2:
a) See Handwritten Page for Marginal Density Formula
 
```{r Problem 2 setup}
# marginal probabilities
prob <- c( .5, .5 )

# conditional means and standard deviations. Set sigma = 2 for both cases
mu <- c(1, 2)
sigma <- c( 2, 2)
```



```{r Problem 2 Marginal Plot}
# number of samples
n.samples <-1000

# simulate genotypes
param <- sample( c( "Equals 1", "Equals 2"), n.samples, replace = T, prob = prob )

# simulate scores conditional on param
score.sim <- data.frame(score = rnorm(n.samples, mu[as.numeric(as.factor(param))], 
               sigma[as.numeric(as.factor(param))]))

# plot histogram of the samples
p <- ggplot(score.sim, aes( x = score)) +
       geom_histogram( aes(y = ..density.. ), fill = "blue", alpha = 0.2)
    

p

# function that calculates the marginal density of scores
marginal.density <- function(x)
                {
                  prob[1] * dnorm( x, mu[1], sigma[1] ) +
                  prob[2] * dnorm( x, mu[2], sigma[2] )
                }

# add marginal density to the histogram
p + stat_function( fun = marginal.density, lwd = 1)


```

b) What is P(θ=1|y=1)P(θ=1|y=1)  when σ=2 ?
```{r Problem 2 Part b}

#calculate P(y=1| param =1)
pnorm(q= 1, mean=1, sd = 2)

#calculate P(y=1| param =2)
pnorm(q= 1, mean=2, sd = 2)
print("When sd is 2: ")
result = (.5*.5)/(.5*.5 + .3085*.5)
result


```

c) Describe how P(θ=1|y=1) changes as you vary σσ.
```{r Problem 2 Part C}

#INCREASE sd

#calculate P(y=1| param =1)
a = pnorm(q= 1, mean=1, sd = 5 )
#calculate P(y=1| param =2)
b = pnorm(q= 1, mean=2, sd = 5 )

result = (a*.5)/(a*.5 + b*.5)
print("When sd is 5 (larger): ")
result


#Decrease sd

#calculate P(y=1| param =1)
a = pnorm(q= 1, mean=1, sd = .1 )
#calculate P(y=1| param =2)
b = pnorm(q= 1, mean=2, sd = .1 )

result = (a*.5)/(a*.5 + b*.5)
print("When sd is .1 (smaller): ")
result


```

When sd gets larger, the width of the distribution increases, making it more likely that y = 1 when θ is 2. Thus, the P(θ=1|y=1) is less. When sd gets smaller, the width of the distribution decreases, making it less likely that y = 1 when θ is 2. Thus, the P(θ=1|y=1) is more.
